R-LIME: LIME法における矩形制約と最適化

### 2024-06-25 paper.tex
build: organize directory structure and refine environment

\documentclass[runningheads]{llncs}

@@

% set the path for figures
\graphicspath{{./data/}{./python/data/}}

@@

% settings for cleveref
\crefname{figure}{Fig.}{Figs.}
\crefname{table}{Table.}{Tables.}

% import the TeX file containing macro definitions
\input{macro}

@@

% Authentic title
\title{%
  R-LIME: Rectangular Constraints and Optimization for Local Interpretable
  Model-agnostic Explanation Methods}

@@

% The abstract should briefly summarize the contents of the paper in
% 150--250 words.
\begin{abstract}
  In recent years, complex machine learning models such as deep neural networks
  have been used in various industrial fields due to their high accuracy.
  However, its complexity has been a major obstacle to implementation
  in decision-making situations where transparency of the decision process is required.
  In order to address this problem,
  various post-hoc explanation methods have been proposed,
  but they have not been able to achieve interpretability of
  both the explanation and its scope.
  We propose a new method, R-LIME,
  which interprets complex classifiers in an interpretable scope.
  R-LIME locally approximizes a complex decision boundary linearly
  in a rectangular region
  and maximizes the region as long as the accuracy of the linear
  classifier is higher than a specified threshold.
  The resulting rectangular region is interpretable for users because it is
  expressed as a conjunction of feature predicates.
  We demonstrate the effectiveness of the proposed method through qualitative
  and quantitative experiments using a real-world dataset.
  Finally, we discuss limitations and future work of the proposed method.
  % 148 words
  \keywords{Interpretable machine learning \and Local surrogate model}
\end{abstract}

\section{Introduction}
Machine learning models, such as deep learning and random forests,
have been widely employed in various industrial applications
due to their significant improvement in accuracy in recent years.
However,
the increasing complexity and black-box nature of these models pose challenges,
particularly in critical decision-making scenarios like healthcare and finance,
where the opacity of decision rationale becomes a major obstacle to
implementation.
Consequently,
there has been extensive research in the field of post hoc explanations
for machine learning models \cite{%
  ribeiro2016why,ribeiro2018anchors,radulovic2023bella,guidotti2018local}.

Existing methods on post hoc explanations categorize into
\emph{model-dependent} and \emph{model-agnostic} methods,
based on their dependence on the model's structure.
Model-dependent methods primarily focus on deep neural networks and
explain the models' behavior using their parameters.
While they allow a direct approach to the model,
applying the same method to models with different structures is often challenging.
On the other hand,
model-agnostic methods do not rely on the models' parameters and utilize only their output.
Although they are applicable to any model, available information is limited.
Model-agnostic methods further classify into
\emph{global} and \emph{local} methods based on their locality in input space.
Global methods aim to explain the model's behavior across the entire input space,
seeking explanations valid for any input.
However, providing global explanations becomes challenging as the model's complexity increases.
Local methods, in contrast, explain the model's output for a specific input
(focal point) in the vicinity of that input.
While local methods offer more simple and accurate explanations than global methods,
the scope of the explanation is limited to the vicinity of the focus point.

This paper specifically focuses on \emph{local} \emph{model-agnostic} methods.
For instance, LIME (Local Interpretable Model-agnostic Explanations)
\cite{ribeiro2016why} approximates the complex model's decision boundary
locally using a simpler model around the given focal point.
Although LIME enables users to interpret the local behavior of complex models,
it does not explicitly provide the effective scope of the explanation
\cite{ribeiro2018anchors}.
This limitation can lead users to overgeneralize the knowledge derived from the explanation.
Without a clear indication of the scope of the explanation,
users might mistakenly apply the derived knowledge to instances outside this scope,
potentially leading to a misunderstanding of the black-box model's behavior.

To address this limitation of LIME,
we propose a new method called R-LIME (Ruled LIME),
which interprets complex classifiers in an interpretable region.
R-LIME locally approximates a complex decision boundary linearly
in a rectangular region and maximizes the region as long as the accuracy of the linear
classifier is higher than a specified threshold.
The resulting rectangular region is interpretable for users because it is
expressed as a conjunction of feature predicates.

\section{Related Work}
\begin{figure}[t]
  \centering
  \vspace{0.5cm}
  \begin{subfigure}[t]{0.55\textwidth}
    \centering
    \includegraphics[width=\textwidth]{lime_vs_anchor_exp_a.png}
    \caption{Instances around the focal point.}\label{fig:lime_vs_anchor_exp_a}
    \vspace{0.5cm}
  \end{subfigure}
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{lime_vs_anchor_exp_b.png}
    \caption{%
      Explanations by LIME\@.
      They provide contributions of each feature to the output,
      but do not explicitly indicate its scope.
    }\label{fig:lime_vs_anchor_exp_b}
    \vspace{0.5cm}
  \end{subfigure}
  \begin{subfigure}[t]{0.55\textwidth}
    \centering
    \includegraphics[width=\textwidth]{lime_vs_anchor_exp_c.png}
    \caption{%
      Explanations by Anchor.
      They provide the effective scope of the explanation,
      but do not provide details about the influence of each feature.
    }\label{fig:lime_vs_anchor_exp_c}
  \end{subfigure}
  \caption{%
    Explanations by LIME and Anchor
    for an LSTM sentiment prediction model \cite{ribeiro2018anchors}.
  }\label{fig:lime_vs_anchor_exp}
\end{figure}
In this chapter,
we provide an overview of existing research on local model-agnostic explanations,
particularly focusing on studies closely related to our proposed method.

\subsection{LIME (Local Interpretable Model-agnostic Explanations) \cite{ribeiro2016why}}
LIME locally
approximates a trained black-box classifier $f: \mathbb{R}^d \to \{0,1\}$
around a focal point $x \in \mathbb{R}^d$
by a linear classifier $g: \mathbb{R}^d \to \{0,1\}$
(\cref{fig:lime}).
The process involves:
\begin{enumerate}
  \item Obtaining a set of perturbed samples $\mathcal{Z}_p$ around $x$
        and the set of pseudo-labels $f(\mathcal{Z}_p) = \{f(z) \mid z \in \mathcal{Z}_p\}$.
  \item Learning a linear classifier $g$
        using $\mathcal{Z}_p$ and $f(\mathcal{Z}_p)$
        by minimizing the following loss function:
        \begin{equation}
          \label{eq:lime_loss}
          \mathcal{L}(f,g,\pi_x)=\sum_{z\in\mathcal{Z}_p}
          \pi_x(z){\left(f(z)-g(z)\right)}^2,
        \end{equation}
        where $\pi_x(z)$ is a weight function designed to be larger for samples
        closer to $x$, typically implemented using an exponential kernel.
\end{enumerate}

LIME provides valuable insights into the local behavior of the model
by showing the contribution of each feature to the output $f(x)$.
However, due to not explicitly indicating the perturbation region,
users cannot assess the effective scope of the explanation
\cite{ribeiro2018anchors}.
An example of explanation by LIME for an LSTM sentiment prediction model
is illustrated in \cref{fig:lime_vs_anchor_exp_b}.
The left explanation suggests that
the word "not" contributes to the model's positive prediction,
which does not apply to the instance on the right.
However, without the explicit perturbation region,
users might mistakenly apply the left explanation to the right instance,
potentially leading to a misunderstanding of the black-box model's behavior
\cite{ribeiro2018anchors}.

\subsection{Anchor \cite{ribeiro2018anchors}}\label{sec:anchor}
Anchor represents a rectangular region containing the focal point $x$,
expressed as a conjunction of feature predicates (a rule),
that maximizes the probability of the black-box classifier $f$
outputting $f(x)$ within the region.
It aims to highlight important features
contributing significantly to the output (\cref{fig:anchor}).

For a discrete $m$-dimensional input space $\ispace$
with a trained black-box classifier $f: \ispace \to \{0,1\}$,
an instance $x \in \ispace$,
and a distribution $\mathcal{D}$ over the input space,
a rule $A(z) = a_{i_1}(z) \wedge a_{i_2}(z) \wedge \dots \wedge a_{i_t}(z)$ is defined.
The predicates $a_i(z)$ evaluate to true ($=1$) when $z_i = x_i$ and false ($=0$) otherwise.
The accuracy $\Prec(A)$ and coverage $\Cov(A)$ of the rule $A$ are defined as follows:
\begin{align}
  \Prec(A) & =\mathbb{E}_{z\sim\mathcal{D}(z|A)}
  [\mathbbm{1}_{f(z)=f(x)}], \label{eq:anchor_def_prec} \\
  \Cov(A)  & =\mathbb{E}_{z\sim\mathcal{D}(z)}[A(z)].
\end{align}

where $\mathcal{D}(z|A)$ is the conditional distribution in the region
where the rule $A$ returns true.
$\Prec(A)$ represents the probability that the output of $f$ matches
between the perturbation $z\sim\mathcal{D}(z|A)$ and the focal point $x$,
and $\Cov(A)$ expresses the probability that the perturbation $z$ fits into $A$.

Anchor maximizes coverage under the constraint that
the accuracy of the rule $A$ exceeds a given threshold $\tau$.
However, \cref{eq:anchor_def_prec} is not directly computable.
Introducing a confidence level $1-\delta$ $(0\le\delta\le1)$,
the accuracy constraint is relaxed as follows:
\begin{equation}
  \label{eq:const_prec}
  P(\Prec(A)\ge\tau)\ge1-\delta.
\end{equation}
Thus, the following optimization problem is solved:
\begin{equation}
  \label{eq:main_problem}
  A^*=\underset{A \text{~s.t. } P(\Prec(A)\ge\tau)\ge1-\delta\wedge A(x)=1}
  {\arg\max}\Cov(A).
\end{equation}

While Anchor provides an effective scope of the explanation,
the information it includes is less than LIME,
limiting its utility \cite{ribeiro2018anchors}.
An example of Anchor explanation for an LSTM sentiment prediction model
is depicted in \cref{fig:lime_vs_anchor_exp_c}.
The left explanation suggests that changing words other than "not" and "bad"
has little impact on the classifier's output.
While it clearly does not apply to the right instance,
the explanation does not provide details about the influence of each word,
resulting in less user insight into the model's behavior compared to LIME
(\cref{fig:lime_vs_anchor_exp_b}).

\subsection{%
  BELLA (Black box model Explanations by Local Linear Approximations)
  \cite{radulovic2023bella}
}
BELLA addresses the non-optimization of the approximation region,
a drawback of LIME,
by learning a local linear classifier $g$
using a subset $Z'$ of the dataset $Z$
instead of perturbed samples (\cref{fig:bella}).
The process involves:
\begin{enumerate}
  \item Computing the distance $d_x(z)$ for all instances $z$ in $Z$
        with respect to the focal point $x$.
  \item Selecting the $k$ instances with the smallest $d_x(z)$,
        forming a subset $Z_s(k) \subseteq Z$.
  \item Learning the local linear classifier $g$ using $Z_s(k)$
        and calculating the similarity measure $\mathcal{R}(k)$ between $f$ and $g$
        over $k=1,2,\dots,|Z|$.
  \item Obtaining the optimal $k^*$ by maximizing $\mathcal{R}(k)$.
\end{enumerate}

BELLA aims to enhance the reliability of explanations
by optimizing the approximation region for high-accuracy linear approximations.
However, it assumes direct access to the dataset $Z$ used for training the classifier,
making it inapplicable in scenarios where direct access to the dataset is restricted.
Additionally,
as BELLA presents the approximation region as a subset of the dataset,
its interpretability is lower compared to Anchor.

\section{Proposed Method}
\begin{figure}[tb]
  \centering
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{lime}
    \caption{%
      LIME: Locally approximates the decision boundary around the focal point.
    }\label{fig:lime}
  \end{subfigure}%
  \hfill
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{anchor}
    \caption{%
      Anchor: Maximizes coverage of a rectangular region
      containing the focal point under accuracy constraints.
    }\label{fig:anchor}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=0.85\textwidth]{bella}
    \caption{%
      BELLA: Explores a subset of the dataset,
      maximizing similarity between the local linear classifier
      and the black-box classifier.
    }\label{fig:bella}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \includegraphics[width=\textwidth]{rlime3}
    \caption{%
      R-LIME:
      Maximizes coverage of a rectangular region containing the focal point
      under lower constraints on the accuracy of the linear classifier.
    }\label{fig:rlime}
  \end{subfigure}
  \caption{%
    Visual comparison of LIME, Anchor, BELLA and R-LIME (our proposed method).
    The solid line represents the rectangular region containing the focal
    point, and the dashed line represents the learned approximation model.
  }
\end{figure}

\subsection{Overview}
We propose R-LIME,
a method that aims to address the limitations of LIME \cite{ribeiro2016why},
Anchor \cite{ribeiro2018anchors} and BELLA \cite{radulovic2023bella}.
Our method locally approximates the given black-box classifier $f$
around the focal point $x$ by a linear classifier $g$ similar to LIME and BELLA,
but samples parturbation from a rectangular region similar to Anchor
(\cref{fig:rlime}).
The accuracy of $g$ is defined as the "accuracy" of the rule,
and the generality of explanations is defined as the "coverage" of the rule.

Anchor maximizes the coverage of region $A$
as long as the probability of the output of the black-box classifier $f$ matching $f(x)$
within $A$ exceeds a given threshold $\tau$.
The proposed method, on the other hand, learns an linear classifier $g$
within the rectangular region $A$ and maximizes the coverage of $A$
under lower constraints on the accuracy of $g$.
We modify the Anchor's definition of accuracy in \cref{eq:anchor_def_prec}
as follows:
\begin{equation}
  \Prec(A)=\underset{g\in G}{\max}\ \mathbb{E}_{z\sim\mathcal{D}(z|A)}
  [\mathbbm{1}_{f(z)=g(z)}]. \label{eq:def_prec}
\end{equation}
where $G$ is the set of possible linear classifiers.
By solving the optimization problem in \cref{eq:main_problem}
under the modified accuracy definition in \cref{eq:def_prec},
we can select the rule that enables explanation with high accuracy and generality.
The characteristics of R-LIME are listed below:
\begin{itemize}
  \item The approximation region is optimal and interpretable.
  \item The dataset is not directly accessed, and only the distribution is utilized.
  \item The number of samples is dynamically determined
        based on the estimated accuracy without the need for predefining it.
\end{itemize}

\subsection{Algorithm}\label{sec:alg}
{%
  \begin{figure}[t]
    \centering
    \includegraphics[width=0.32\textwidth]{rlime1}
    \includegraphics[width=0.32\textwidth]{rlime2}
    \includegraphics[width=0.32\textwidth]{rlime3}
    \caption{%
      Overview of the R-LIME algorithm.
      The progression of the algorithm is illustrated from left to right.
      The solid line represents the rectangular region $A$,
      and the dashed line represents the linear approximation model $g$
      learned within $A$.
      The initial value of $A$ is an empty rule (entire input space),
      and predicates are added to $A$, reducing coverage.
      The process continues until $\Prec(A)\ge\tau$,
      at which point the rule with the maximum coverage is output.
    }
  \end{figure}
  \begin{algorithm}[p]
    \input{alg1}
  \end{algorithm}
  \begin{algorithm}[p]
    \input{alg2}
  \end{algorithm}
  \def\myidt{\hspace{\algorithmicindent}}
  \begin{algorithm}[p]
    \input{alg3}
  \end{algorithm}
  \begin{algorithm}[p]
    \input{alg4}
  \end{algorithm}
}

Our method's algorithm is mainly based on that used in Anchor.
For non-convex optimization problems like \cref{eq:main_problem},
greedy search are often used.
However,
greedy methods tend to converge to local optima, and to address this,
R-LIME utilizes beam search,
which selects multiple candidates at each iteration.
The pseudocode is shown in \cref{alg:greedy_search}.

\subsubsection{Generation of New Candidate Rules}
To generate new candidate rules,
one additional predicate is added to each of the $B$ candidate rules
selected in the previous iteration.
The pseudocode is shown in \cref{alg:generate_cands}.
$T(x)$ is the set of attribute-value pairs (predicates) that are true for $x$,
and $T(x)\setminus A$ is the set of predicates in $T(x)$ not included in rule $A$.

>>>>

\def\mode{1} % 0: Graduation thesis, 1: ICPR paper

\ifnum\mode=0
  \documentclass[11pt]{article}
  \usepackage{luatexja}
  \usepackage{src/style/prmlthesis}
  \usepackage[a4paper]{geometry}
\else
  \documentclass[runningheads]{llncs}
\fi

@@

% set the path for figures
\graphicspath{{./src/}{../python/data/}{src/experiments/}}

@@

% settings for cleveref
\crefname{table}{Table}{Tables}
\crefname{algorithm}{Algorithm}{Algorithms}
\ifnum\mode=0
  \crefname{figure}{Figure}{Figures}
\else
  \crefname{figure}{Fig.}{Figs.}
\fi

% import the TeX file containing macro definitions
\input{src/tex/macro}

@@

\ifnum\mode=0
  \title{%
    R-LIME: LIME法における矩形制約と最適化
  }
  \etitle{%
    R-LIME: Rectangular Constraints and Optimization for Local Interpretable
    Model-agnostic Explanation Methods
  }
  \author{大原 玄嗣 (Genji Ohara)}
  \institute{%
    北海道大学 工学部 \\
    情報エレクトロニクス学科 \\
    情報理工学コース \\
    情報認識学研究室
  }
  \date{2024年3月}
  % \institute{%
  %   Laboratory for Pattern Recognition and Machine Learning,\\
  %   Division of Computer Science and Information Technology,\\
  %   Department of Electronics and Information Engineering,\\
  %   School of Engineering, Hokkaido University
  % }
\else
  % The title of the paper
  \title{%
    R-LIME: Rectangular Constraints and Optimization for Local Interpretable
    Model-agnostic Explanation Methods
  }

@@

\ifnum\mode=0
  \pagestyle{plain}
  \pagenumbering{roman}
  \tableofcontents
  \newpage
  \listoffigures\listoftables\clearpage
  \AddToHook{cmd/section/before}{\clearpage}
  \pagestyle{plain}
  \pagenumbering{arabic}
\else
  % The abstract should briefly summarize the contents of the paper in
  % 150--250 words.
  \begin{abstract}
    In recent years,
    complex machine learning models have been introduced
    in various industrial fields due to their high accuracy.
    However,
    their increasing complexity has been a major obstacle to implementation
    in sensitive decision-making situations.
    In order to address this problem,
    various post-hoc explanation methods have been proposed,
    but they have not been able to achieve interpretability of
    both the explanation and its scope.
    We propose a new method, R-LIME,
    which interprets a complex classifier in an interpretable scope.
    R-LIME locally and linearly approximates a complex decision boundary
    of a black-box classifier in a rectangular region
    and maximizes the region as long as the approximation accuracy
    exceeds a given threshold.
    The resulting rectangular region is interpretable for users because it is
    expressed as a conjunction of feature predicates.
    Through qualitative and quantitative comparisons with the existing method
    on a real-world dataset,
    we demonstrate that R-LIME provides more reliable and interpretable
    explanations than existing methods.
    % 153 words
    \keywords{Interpretable machine learning \and Local surrogate model}
  \end{abstract}
\fi

\section{Introduction}
In recent years,
complex machine learning models, such as deep neural networks and random forests,
have been widely introduced in various industrial fields
due to their significant improvement in accuracy.
However,
their increasing complexity and black-box nature pose challenges,
particularly in critical decision-making scenarios such as healthcare and finance,
where the opacity of decision process becomes a major obstacle to implementation.
In order to address this problem,
there has been extensive research in the field of post-hoc explanations
for machine learning models~\cite{%
  guidotti2018local,radulovic2023bella,ribeiro2016why,ribeiro2018anchors}.
Existing post-hoc explanation methods are categorized
into \emph{model-dependent} and \emph{model-agnostic} methods
based on their dependence on the model's structure,
and the latter are further classified into \emph{global} and \emph{local} methods
based on their locality in input space~\cite{samek2021explaining}.
\begin{figure}[tbp]
  \def\scale{0.38}
  \centering
  \begin{subfigure}[t]{0.55\textwidth}
    \centering
    \includegraphics[scale=\scale]{example-instance}
    \caption{Focal point.}\label{fig:example-instance}
    \vspace{0.5cm}
  \end{subfigure}
  \begin{subfigure}[t]{0.45\textwidth}
    \begin{subfigure}[t]{\textwidth}
      \centering
      \includegraphics[scale=\scale]{example-lime}
      \caption{%
        Explanation by LIME\@.
        It provides contributions of each feature to the output,
        but does not explicitly indicate its scope.
      }\label{fig:example-lime}
      \vspace{0.4cm}
    \end{subfigure}
    \begin{subfigure}[t]{\textwidth}
      \centering
      \includegraphics[scale=\scale]{example-anchor}
      \caption{%
        Explanation by Anchor.
        It provides the effective scope of the explanation,
        but does not show the influence of each feature.
      }\label{fig:example-anchor}
    \end{subfigure}
  \end{subfigure}
  \hspace{0.3cm}
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \ifnum\mode=0
      \vspace{-1.79cm}
    \else
      \vspace{-2.24cm}
    \fi
    \includegraphics[scale=\scale]{example-rlime}
    \caption{%
      Explanation by R-LIME\@.
      It provides both contributions of each feature and
      its effective scope.
    }\label{fig:example-rlime}
  \end{subfigure}
  \caption[Example of explanations by LIME, Anchor and R-LIME]{%
    Example of explanations by LIME~\cite{ribeiro2016why},
    Anchor~\cite{ribeiro2018anchors} and R-LIME (our proposed method)
    for a sentiment prediction model.
  }\label{fig:example}
\end{figure}

In this paper,
we focus on \emph{local} and \emph{model-agnostic} methods.
LIME~\cite{ribeiro2016why} and Anchor~\cite{ribeiro2018anchors}
are representative local model-agnostic methods.
An example of explanations by LIME and Anchor
for a sentiment prediction model is illustrated in \cref{fig:example}.
LIME linearly approximates the complex decision boundary
around the given focal point,
then provides the weights of the linear model as the contribution of each feature
to the output.
The explanation by LIME (\cref{fig:example-lime}) suggests that
the word ``not'' mainly contributes to the positive prediction,
but does not explicitly indicate its effective scope.
Without the scope,
users might mistakenly apply the knowledge derived from the explanation
to other instances far from the focal point,
potentially leading to misunderstanding of the black-box model's behavior
\cite{ribeiro2018anchors}.
For this example,
users may apply the derived insights
to the sentence ``This book is not good.''
and mistakenly conclude that the word ``not''
mainly contributes to the positive prediction for this sentence as well,
which is obviously incorrect.
Anchor maximizes the coverage of a rectangular region containing the focal point
as long as the probability of the black-box classifier outputting
the same label as the focal point within the region exceeds a given threshold.
While Anchor provides an effective scope of the explanation,
users can get less insight compared to LIME\@.
The explanation by Anchor (\cref{fig:example-anchor})
suggests that replacing words other than ``not'' and ``bad''
has little impact on the classifier's output.
While it clearly cannot be applied to the sentence ``This book is not good''
because of not including the word ``bad'',
the explanation does not provide details about the influence of each word,
resulting in less user insight into the model's behavior compared to LIME.

To address these limitations,
we propose a new method called R-LIME (Ruled LIME),
which provides both the contributions of each feature to the output
and the effective scope of the explanation.
R-LIME linearly approximates a complex decision boundary
in a rectangular region and maximizes the region
as long as the accuracy of the linear classifier exceeds a given threshold.
The region is interpretable for users because it is
expressed as a conjunction of feature predicates.
An example of the explanation by R-LIME for a sentiment prediction model
is shown in \cref{fig:example-rlime}.
It is clear that users can apply the insights derived from the explanation
only to the sentences containing the word ``not''.

\section{Related Work}
\begin{figure}[tbp]
  \centering
  \def\w{2.5}
  \def\ww{5.0}
  \def\h{0.5}
  \def\hh{1.3}
  \def\hhh{1.8}
  \def\hhhh{2.6}
  \begin{tikzpicture}[scale=0.9,nodestyle/.style={rectangle,auto}]
    \node[nodestyle] (post-hoc) at (0,0){post-hoc};
    \node[nodestyle] (dependent) at (-\w,-\hh){%
      \begin{tabular}{c}
        model-dependent \\
        (DTD~\cite{montavon2017explaining}, LRP~\cite{bach2015pixel})
      \end{tabular}
    };
    \node[nodestyle] (agnostic) at (\w,-\hh){\textbf{model-agnostic}};
    \node[nodestyle] (global) at (0,-\hhhh){%
      \begin{tabular}{c}
        global \\
        (PDP~\cite{friedman2001greedy}, ALE~\cite{apley2020visualizing})
      \end{tabular}
    };
    \node[nodestyle] (local) at (\ww,-\hhhh){%
      \begin{tabular}{c}
        \textbf{local} \\
        (LIME~\cite{ribeiro2016why}, Anchor~\cite{ribeiro2018anchors}, \underline{R-LIME})
      \end{tabular}
    };
    \draw(post-hoc) -- (0,-\h);
    \draw(dependent) -- (-\w,-\h) -- (\w,-\h) -- (agnostic);
    \draw(agnostic) -- (\w,-\hhh);
    \draw(global) -- (0,-\hhh)  -- (\ww,-\hhh) -- (local);
  \end{tikzpicture}
  \caption{%
    Categorization of post-hoc explanation methods.
    We focus on \emph{model-agnostic} and \emph{local} methods,
    which explain model's local behavior using only its output.
  }\label{fig:post-hoc}
\end{figure}
In this section,
we overview existing research on post-hoc explanation methods,
which explain the behavior of black-box models already trained.
As shown in \cref{fig:post-hoc},
post-hoc methods are classified into several categories.

They are broadly divided into
\emph{model-dependent} and \emph{model-agnostic} methods
based on their dependence on the model's structure.
Model-dependent methods,
such as deep Taylor decomposition (DTD)~\cite{montavon2017explaining}
and layer-wise relevance propagation (LRP)~\cite{bach2015pixel},
most of which focus on neural networks and
explain the model's behavior using its parameters~\cite{samek2021explaining}.
While these methods provide detailed explanations
(e.g., layer-wise explanations for neural networks),
it is often challenging
to apply the same method to models with different structures.
On the other hand,
model-agnostic methods use only the output of the model.
Although they are applicable to any model,
they cannot explain the reasoning process inside the model.

Furthermore, model-agnostic methods are categorized into
\emph{global} and \emph{local} methods based on their locality in input space.
Global methods,
such as partial dependence plots (PDP)~\cite{friedman2001greedy}
and accumulated local effects (ALE)~\cite{apley2020visualizing},
aim to explain the model's behavior across the entire input space.
However, providing global explanations becomes challenging
as the model's complexity increases.
In contrast, local methods,
such as local interpretable model-agnostic explanations (LIME)
\cite{ribeiro2016why}
and Anchor~\cite{ribeiro2018anchors},
explain model's behavior in the vicinity of a specific input.
While they offer explanations more simple and accurate
than global methods,
the scope of the explanation is limited locally.

\section{Proposed Method}
\begin{figure}[tbp]
  \centering
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{visual-lime}
    \caption{%
      LIME: Locally approximates the decision boundary around the focal point.
    }\label{fig:lime}
  \end{subfigure}%
  \hspace{0.03\textwidth}
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{visual-anchor}
    \caption{%
      Anchor: Maximizes coverage of a rectangular region
      containing the focal point under accuracy constraints.
    }\label{fig:anchor}
  \end{subfigure}
  \hspace{0.03\textwidth}
  \begin{subfigure}[t]{0.3\textwidth}
    \centering
    \includegraphics[width=\textwidth]{visual-rlime3}
    \caption{%
      R-LIME:
      Maximizes coverage of a rectangular region
      under lower constraints on the accuracy of the linear classifier.
    }\label{fig:rlime}
  \end{subfigure}
  \caption[Visual comparison of LIME, Anchor and R-LIME]{%
    Visual comparison of LIME, Anchor and R-LIME (our method).
    The dashed line represents the local linear approximation model,
    and the solid line represents the rectangular region containing the focal point.
  }
\end{figure}

\subsection{Previous Work}
We specifically focus on \emph{local} and \emph{model-agnostic} methods.
In this section,
we briefly review existing research on local model-agnostic explanations,
particularly focusing on studies closely related to our proposed method.

\subsubsection[LIME]{%
  LIME (Local Interpretable Model-agnostic Explanations)
  \cite{ribeiro2016why}
}
LIME locally
approximates a black-box classifier $f: \mathbb{R}^m \to \{0,1\}$
around a focal point $x \in \mathbb{R}^m$
by a linear classifier $g: \mathbb{R}^m \to \{0,1\}$
(\cref{fig:lime}).
The approximation is performed by the following steps:
\begin{enumerate}
  \item Generating a set of perturbed samples $\mathcal{Z}_p$ around $x$
        and the set of pseudo-labels $f(\mathcal{Z}_p) = \{f(z) \mid z \in \mathcal{Z}_p\}$.
        (i) $x$ is converted into a binary vector
        $x'\in\{0,1\}^{m'}$,
        (ii) perturbed samples are generated by drawing non-zero elements
        from $x'$ uniformly at random,
        and (iii) the perturbed samples are converted back to the original space.
  \item Learning a linear classifier $g$
        using $\mathcal{Z}_p$ and $f(\mathcal{Z}_p)$
        by minimizing the following loss function:
        \begin{equation}
          \label{eq:lime_loss}
          \mathcal{L}(f,g,\pi_x)=\sum_{z\in\mathcal{Z}_p}
          \pi_x(z){\left(f(z)-g(z)\right)}^2,
        \end{equation}
        where $\pi_x(z)$ is a weight function designed to be larger for samples
        closer to $x$, typically defined using an exponential kernel.
\end{enumerate}
LIME provides valuable insights into the local behavior of the model
by showing the contribution of each feature to the output $f(x)$.
However, it does not explicitly indicate the region for generating perturbed samples,
making it difficult for users to assess the effective scope of the explanation
\cite{ribeiro2018anchors}.

\subsubsection[Anchor]{%
  Anchor~\cite{ribeiro2018anchors}
}\label{sec:anchor}
Anchor maximizes the coverage of a rectangular region
containing the focal point $x$,
expressed as a conjunction of feature predicates (a rule)
as long as the probability of the black-box classifier $f$
outputting $f(x)$ within the region exceeds a given threshold $\tau$
(\cref{fig:anchor}).
It aims to highlight important features
contributing significantly to the output.
For a discrete $m$-dimensional input space $\ispace$,
a trained black-box classifier $f: \ispace \to \{0,1\}$,
an instance $x \in \ispace$
and a distribution $\mathcal{D}$ over the input space,
a rule $A(z) = a_{i_1}(z) \wedge a_{i_2}(z) \wedge \dots \wedge a_{i_t}(z)$ is defined.
The predicate $a_i(z)$ evaluates to true ($=1$) when $z_i = x_i$ and false ($=0$) otherwise.
The reliability of the explanation is defined as the ``accuracy'' of the rule,
and the generality of the explanation is defined as the ``coverage'' of the rule.
The accuracy $\Prec(A)$ and coverage $\Cov(A)$ of the rule $A$ are defined as follows:
\begin{align}
  \Prec(A) & =\mathbb{E}_{z\sim\mathcal{D}(z|A)}
  [\mathbbm{1}_{f(z)=f(x)}], \label{eq:anchor_def_prec} \\
  \Cov(A)  & =\mathbb{E}_{z\sim\mathcal{D}(z)}[A(z)],
\end{align}
where $\mathcal{D}(z|A)$ is the conditional distribution in the region
where the rule $A$ returns true.
$\Prec(A)$ represents the probability that the output of $f$ matches
between the perturbation $z\sim\mathcal{D}(z|A)$ and the focal point $x$,
and $\Cov(A)$ expresses the probability that the perturbation $z$ fits into $A$.
Anchor maximizes coverage as long as
the accuracy of the rule $A$ exceeds a given threshold $\tau$.
However, \cref{eq:anchor_def_prec} is not directly computable.
Introducing a confidence level $1-\delta$ $(0\le\delta\le1)$,
the accuracy constraint is relaxed as follows:
\begin{equation}
  \label{eq:const_prec}
  P(\Prec(A)\ge\tau)\ge1-\delta.
\end{equation}
Thus, the following optimization problem is solved:
\begin{equation}
  \label{eq:main_problem}
  A^*=\underset{A \text{~s.t. } P(\Prec(A)\ge\tau)\ge1-\delta\wedge A(x)=1}
  {\arg\max}\Cov(A).
\end{equation}

\subsection{Overview}
We propose R-LIME,
a method that aims to address the limitations of LIME~\cite{ribeiro2016why}
and Anchor~\cite{ribeiro2018anchors}.
Our method locally approximates the given black-box classifier $f$
around the focal point $x$ by a linear classifier $g$ similar to LIME,
but generates the perturbed samples
from a rectangular region similar to Anchor
so that the generality of approximation is explicitly provided
(\cref{fig:rlime}).

Anchor maximizes the coverage of region $A$
as long as the probability of the output of the black-box classifier $f$ matching $f(x)$
within $A$ exceeds a given threshold $\tau$.
R-LIME, on the other hand, learns a linear classifier $g$
within the rectangular region $A$ and maximizes the coverage of $A$
under lower constraints on the accuracy of $g$.
We modify Anchor's definition of accuracy in \cref{eq:anchor_def_prec}
as follows:
\begin{equation}
  \Prec(A)=\underset{g\in G}{\max}\ \mathbb{E}_{z\sim\mathcal{D}(z|A)}
  [\mathbbm{1}_{f(z)=g(z)}], \label{eq:def_prec}
\end{equation}
where $G$ is a hypothesis space of possible linear classifiers.
By solving the optimization problem in \cref{eq:main_problem}
under the modified definition of accuracy in \cref{eq:def_prec},
we can select the rule that enables explanation with high accuracy and generality.

\subsection{Algorithm}\label{sec:alg}
{%
  \begin{figure}[t]
    \centering
    \includegraphics[width=0.32\textwidth]{visual-rlime1}
    \includegraphics[width=0.32\textwidth]{visual-rlime2}
    \includegraphics[width=0.32\textwidth]{visual-rlime3}
    \caption[Overview of the R-LIME algorithm]{%
      Overview of the R-LIME algorithm.
      The progression of the algorithm is illustrated from left to right.
      The solid line represents the rectangular region $A$,
      and the dashed line represents the linear approximation model $g$
      learned within $A$.
      The initial value of $A$ is an empty rule (entire input space),
      and predicates are added to $A$, reducing coverage.
      The process continues until $\Prec(A)\ge\tau$ is satisfied,
      at which point the rule with the maximum coverage is output.
    }
  \end{figure}
  \def\myidt{\hspace{\algorithmicindent}}
  \ifnum\mode=0
    \begin{algorithm}[t]
      \small
      \input{src/tex/alg1}
    \end{algorithm}
  \else
    \begin{algorithm}[p]
      \small
      \input{src/tex/alg1}
    \end{algorithm}
  \fi

  \foreach \alg in {2,3,4}{%
      \begin{algorithm}[p]
        \small
        \input{src/tex/alg\alg}
      \end{algorithm}
    }
}
The algorithm of R-LIME
is mainly based on that used in Anchor\cite{ribeiro2018anchors}.
For non-convex optimization problems like \cref{eq:main_problem},
greedy search are often used.
But greedy methods tend to converge to local optima,
so we use beam search, which selects multiple candidates at each iteration.
The pseudo-code is shown in \cref{alg:greedy_search}.
